Location and Scale-Invariant Power Transformations for Transforming Data to

### 2025-12-19 manuscript_rmarkdown/manuscript.Rmd
Rework text.

\begin{definition}[Box-Cox power transformation]
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a finite sequence of length $n > 0$.
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
Then the Box-Cox power transformation of element $x_i$ is defined as:
\begin{equation}
\label{eqn:box-cox-original}
\phi_{\text{BC}}^\lambda (x_i) = 
\begin{cases}
\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(x_i) & \text{if } \lambda = 0
\end{cases}
\end{equation}
\end{definition}

@@

\begin{definition}[Yeo-Johnson power transformation]
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ be a finite sequence of length $n > 0$.
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
\begin{equation}
\label{eqn:yeo-johnson-original}
\phi_{\text{YJ}}^\lambda (x_i) = 
\begin{cases}
\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
\end{cases}
\end{equation}
\end{definition}

@@

\begin{definition}[Numeric feature]
A numeric feature is a finite sequence $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
\end{definition}

@@

\begin{definition}[Location- and scale- invariant Box-Cox power transformation]
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i - x_0 > 0 \right\}$ be a finite sequence of length $n > 0$.

@@

\begin{definition}[Location- and scale- invariant Yeo-Johnson power transformation]
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathcal{R} \right\}$ be a finite sequence of length $n > 0$.

@@

\begin{definition} [Log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.

@@

\begin{definition}[Log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.

@@

\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
Let $\mu_w$ be the weighted mean of the Box-Cox transformed sequence:
\begin{equation*}
\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation*}

Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed sequence:
\begin{equation*}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
\end{equation*}

@@

Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed sequence:
\begin{equation*}
\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation*}

Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed sequence:
\begin{equation*}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
\end{equation*}

@@

\begin{definition}[Central portion of a sequence]
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
percentile values are replaced by the average in their group, i.e.
if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
Furthermore, let central portion $\kappa \in (0,1)$.

@@

\begin{definition}[Residual errors of the central portion of a sequence]

@@

A central normal distribution is any distribution whose quantile function is 
defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
with $\mathcal{N}$ any normal distribution.

@@

For $\kappa = 1.0$, the central normal distribution is a normal distribution.

@@

By the Glivenko-Cantelli theorem, if a sequence $\mathbf{X}$ was sampled from a
central normal distribution, then almost surely
$\lim_{n \rightarrow \infty} \sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$,

@@

\begin{definition}[Central normality test]
The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
from a population that is distributed according to a central normal distribution.

The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
population that is not distributed according to a central normal distribution.
\end{definition}

@@

To assess whether the proposed power transformations lead to values of $\lambda$ 
that are invariant to location and scale of the distribution, we simulated three different sequences.
We first randomly drew $10000$ values from a normal distribution: 
$\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, 
or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution was a right-skewed generalised normal distribution 
$\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution was a left-skewed generalised normal distribution 
$\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
We then computed transformation parameter $\lambda$ using the original definitions (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.

@@

The critical test statistic for the empirical central normality test depends on 
the significance level $\alpha$, sequence length $n$ and central portion $\kappa$.
Critical test statistic values were set through simulation, as described below.

For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
we randomly sampled $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
For each sequence we then computed $\tau_{\kappa}$ for 
$\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.

@@

To assess type I error rates for the empirical central normality test and compare
these to the Shapiro-Wilk test for normality, we randomly drew another 10000 sequences from a
central normal distribution for each 
$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.

Each sequence was created as follows: first $n$ values are sampled from $\mathcal{N}(0, 1)$.
Then, the elements outside the central portion of the sequence (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
Elements with $p_i < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
and elements with $p_i > \frac{1 + \kappa}{2}$ are replaced with values sampled from $\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)$.

For each sequence, the p-value for the respective test was used to reject the null
hypothesis that sequences were derived from a population with a (central) normal distribution.
The null hypothesis was rejected if $p \leq 0.05$.
The type I error rate was determined by computing the fraction of sequences rejected this way.
Figure \ref{fig:empirical-central-normality-error_rate} shows that when the
central portion $\kappa$ of the central normal distribution is equal or greater
than the $\kappa$ parameter for the empirical central normality test, the tests 
are accurate. If $\kappa$ of the empirical central normality test exceeds that 
the central portion $\kappa$ for the underlying distribution, 
tests are conservative. The Shapiro-Wilk test behaves similar to the empirical
central normality test with $\kappa = 1.0$, which is expected. For $n < 10$,
the empirical central normality test can be optimistic.

@@

To determine the weighting function parameters for each of the nine combinations,
$m_d=500$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.
Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
These values were set according to @Tukey1977-xm, as follows.
Let $x^{*} \sim U\left(-2, 2\right)$.
Then the corresponding outlier value was:

@@

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised a composite loss $L = \sum_{i=1}^{m_d} L_{\text{cn},i} + 0.1 \sum_{i=1}^{m_d} L_{\lambda,i}$.
The composite loss consisted of two components: a loss term $L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}$, i.e. the mean of residual 
errors of the central 80% of elements of each sequence $\mathbf{X}_i$, which aimed at optimising central normality; 
and a loss term $L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)$,
with $\lambda_{0,i}$ and $\lambda_{i}$ transformation parameters found
for sequence $\mathbf{X}_i$ prior to and after adding outliers, respectively, 
and tolerance parameter $\xi = 0.5$ for Box-Cox and $\xi = 0.3$ for Yeo-Johnson power transformations.
This second term aimed to prevent solutions that provide small improvements in central normality
at the cost of a poor fit of the tails of sequences.

@@

Three datasets with 100000 sequences each were created to assess transformation
to normality. For each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
elements were randomly drawn, with $\gamma \sim U\left(1, 4\right)$, 
resulting in sequences with between $10$ and $10000$ elements.

- A \textit{clean} dataset with each sequence drawn from $\mathcal{N}(0, 1)$ and
  transformed using an inverse power transformation with randomly drawn 
  transformation parameter $\lambda \sim U\left(0.00, 2.00 \right)$. 
  $\left(\phi_{\text{BC}}^\lambda\right)^{-1}$ and $\left(\phi_{\text{YJ}}^\lambda\right)^{-1}$
  were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
  respectively. Prior to inverse transformation, sequences for Box-Cox transformations
  were shifted into the positive domain, with minimum value $1$.

- A \textit{dirty} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)$,
  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
  The sequences are drawn from a range of skewed, unimodal, but not necessarily 
  normal, distributions, to reflect less \textit{clean} data than presented in the first dataset.

- A \textit{shifted} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)$,
  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
  The generating process creates sequences similar to that in the \textit{dirty}
  dataset, but shifted to a location that leads to normalisation issues observed
  for conventional power transformation methods, as shown in Figure \ref{fig:decreased-normality}.

A dataset with outliers was created for each of the above datasets by replacing
10 percent of elements in each sequence, as described earlier.

@@

1. z-standardisation: $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of sequence $\mathbf{X}$.

@@

This results in nine types of power transformation. Transformation parameters
were optimised for each sequence. Subsequently the sum of residual errors 
and sum of residual errors of the central portion ($\kappa = 0.80$) were computed
for each sequence after transformation. Then, each method was ranked according 
to the sum of residual errors (data without outliers) or the sum of residual
errors of the central portion (data with outliers). The average rank of each
transformation method was computed over all 100000 sequences in each dataset.
Average ranks for Yeo-Johnson transformations are shown in Table
\ref{tab:comparison_methods_simulations_yeo_johnson}.
Location and shift-invariant Yeo-Johnson transformation ranked best for
all datasets without outliers. The robust variant ranked best in dirty and shifted
datasets with outliers, but not in the clean dataset.
Results for Box-Cox transformations are shown in Table \ref{tab:comparison_methods_simulations_box_cox}.

@@

\begin{table}
\begin{center}
\caption{
Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 1000.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations_yeo_johnson}
\begin{tabular}{l | l r r r r r r}

>>>>

\begin{definition}[Box-Cox power transformation]
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a vector of length $n > 0$.
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
Then the Box-Cox power transformation of element $x_i$ is defined as:
\begin{equation}
\label{eqn:box-cox-original}
\phi_{\text{BC}}^\lambda (x_i) = 
\begin{cases}
\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(x_i) & \text{if } \lambda = 0
\end{cases}
\end{equation}
\end{definition}

@@

\begin{definition}[Yeo-Johnson power transformation]
Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ be a vector of length $n > 0$.
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
\begin{equation}
\label{eqn:yeo-johnson-original}
\phi_{\text{YJ}}^\lambda (x_i) = 
\begin{cases}
\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
\end{cases}
\end{equation}
\end{definition}

@@

\begin{definition}[Numeric feature]
A numeric feature is a vector $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
\end{definition}

@@

\begin{definition}[Location- and scale- invariant Box-Cox power transformation]
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
Let $\mathbf{X}$ be a numeric feature of length $n > 0$.

@@

\begin{definition}[Location- and scale- invariant Yeo-Johnson power transformation]
Let $\lambda \in \mathcal{R}$ be a transformation parameter.
Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
Let $\mathbf{X}$ be a numeric feature of length $n > 0$.

@@

\begin{definition} [Log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.

@@

\begin{definition}[Log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.

@@

\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
Let $\mu_w$ be the weighted mean of the Box-Cox transformed feature:
\begin{equation*}
\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation*}

Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed feature:
\begin{equation*}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
\end{equation*}

@@

Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed feature:
\begin{equation*}
\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation*}

Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed feature:
\begin{equation*}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
\end{equation*}

@@

\begin{definition}[Central portion of a numeric feature]
Let $\mathbf{X}$ be a numeric feature of length $n > 0$, as before,
ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
percentile values are replaced by the average in their group, i.e.
if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
Furthermore, let central portion $\kappa \in (0,1)$.

@@

\begin{definition}[Residual errors of the central portion of a feature]

@@

A central normal distribution is any distribution whose quantile function 
satisfies $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
with $\mathcal{N}$ any normal distribution.

@@

A normal distribution is a central normal distribution.

@@

For $\kappa = 1.0$, a central normal distribution is a normal distribution.

@@

If $\mathbf{X}$ is randomly sampled from a normal distribution,
$\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0$, almost surely.
That is, every residual of the central portion of $\mathbf{X}$ will asymptotically
converge to 0.

The latter corollary follows from the definition of the residuals and the restriction
to normal distributions. Residuals were defined using 
$\frac{x_i - \mu_M}{\sigma_M} \approx \hat{F}^{-1}_{\mathcal{N}}(p_i)$, with
$\hat{F}^{-1}_{\mathcal{N}}(p_i)$ the empirical quantile function resembling
the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
Since $\hat{F}^{-1}_{\mathcal{N}}(p) \rightarrow F^{-1}_{\mathcal{N}}(p)$ with
$\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$ [@van-der-Vaart1998-xu],
$\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0$, if and only if
for all $x_i \in \mathbf{X}_c$, $\frac{x_i - \mu_M}{\sigma_M} \rightarrow \hat{F}^{-1}_{\mathcal{N}}(p_i)$.
Thus, the requirement is that the estimators for location and scale $\mu_M$ and $\sigma_M$
are bias-free. M-estimators for location and scale satisfy this condition, if $\mathbf{X}$ 
follows a normal distribution without contamination [Huber1964-xk; @Huber1981-su].

\begin{definition}[Central normality test]
The null-hypothesis $\mathcal{H}_0$ is that $\mathbf{X}$ was randomly sampled
from a population that is distributed according to a central normal distribution.

The alternative hypothesis is that $\mathbf{X}$ was randomly sampled from a
population that is not distributed according to a central normal distribution.
\end{definition}

@@

To assess whether the proposed power transformations lead to values of $\lambda$ 
that are invariant to location and scale of the distribution, we simulated three different numeric features.
We first randomly drew $10000$ values from a normal distribution: 
$\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, 
or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution was a right-skewed generalised normal distribution 
$\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution was a left-skewed generalised normal distribution 
$\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
We then computed transformation parameter $\lambda$ using the original definitions (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.

@@

The critical test statistic for the empirical central normality test depends on 
the significance level $\alpha$, feature length $n$ and central portion $\kappa$.
Critical test statistic values were set through simulation, as described below.

For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
we randomly sampled $m_d = 30000$ ordered numeric features $\mathbf{X}$ from $\mathcal{N}(0,1)$.
For each feature we then computed $\tau_{\kappa}$ for 
$\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.

@@

To assess type I error rates for the empirical central normality test and compare
these to the Shapiro-Wilk test for normality, we randomly drew another 10000 numeric features from a
central normal distribution for each 
$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.

Each feature was created as follows: first $n$ values are sampled from $\mathcal{N}(0, 1)$.
Then, the elements outside the central portion of the feature (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
Elements with $p_i < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
and elements with $p_i > \frac{1 + \kappa}{2}$ are replaced with values sampled from $\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)$.

For each feature, the p-value for the respective test was used to reject the null
hypothesis that features were derived from a population with a (central) normal distribution.
The null hypothesis was rejected if $p \leq 0.05$.
The type I error rate was determined by computing the fraction of features rejected this way.
Figure \ref{fig:empirical-central-normality-error_rate} shows that when the
central portion $\kappa$ of the central normal distribution is equal or greater
than the $\kappa$ parameter for the empirical central normality test, the tests 
are accurate. If $\kappa$ of the empirical central normality test exceeds that 
the central portion $\kappa$ for the underlying distribution, 
tests are conservative. The Shapiro-Wilk test behaves similar to the empirical
central normality test with $\kappa = 1.0$, which is expected. For $n < 10$,
the empirical central normality test can be optimistic.

@@

To determine the weighting function parameters for each of the nine combinations,
$m_d=500$ numeric features $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each feature $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of features with between $50$ and $1000$ elements each.
Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
These values were set according to @Tukey1977-xm, as follows.
Let $x^{*} \sim U\left(-2, 2\right)$.
Then the corresponding outlier value was:

@@

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised a composite loss $L = \sum_{i=1}^{m_d} L_{\text{cn},i} + 0.1 \sum_{i=1}^{m_d} L_{\lambda,i}$.
The composite loss consisted of two components: a loss term $L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}$, i.e. the mean of residual 
errors of the central 80% of elements of each feature $\mathbf{X}_i$, which aimed at optimising central normality; 
and a loss term $L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)$,
with $\lambda_{0,i}$ and $\lambda_{i}$ transformation parameters found
for feature $\mathbf{X}_i$ prior to and after adding outliers, respectively, 
and tolerance parameter $\xi = 0.5$ for Box-Cox and $\xi = 0.3$ for Yeo-Johnson power transformations.
This second term aimed to prevent solutions that provide small improvements in central normality
at the cost of a poor fit of the tails of features.

@@

Three datasets with 100000 numeric features each were created to assess transformation
to normality. For each feature $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
elements were randomly drawn, with $\gamma \sim U\left(1, 4\right)$, 
resulting in features with between $10$ and $10000$ elements.

- A \textit{clean} dataset with each feature drawn from $\mathcal{N}(0, 1)$ and
  transformed using an inverse power transformation with randomly drawn 
  transformation parameter $\lambda \sim U\left(0.00, 2.00 \right)$. 
  $\left(\phi_{\text{BC}}^\lambda\right)^{-1}$ and $\left(\phi_{\text{YJ}}^\lambda\right)^{-1}$
  were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
  respectively. Prior to inverse transformation, features for Box-Cox transformations
  were shifted into the positive domain, with minimum value $1$.

- A \textit{dirty} dataset with each feature drawn from $\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)$,
  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
  The features are drawn from a range of asymmetric unimodal distributions 
  to reflect a wider range of generating distributions than for the 
  data presented in the first dataset.

- A \textit{shifted} dataset with each feature drawn from
  $\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)$,
  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
  The generating process creates features similar to those in the \textit{dirty}
  dataset, but shifted to a location that leads to normalisation issues observed
  for conventional power transformation methods, as shown in Figure \ref{fig:decreased-normality}.

A dataset with outliers was created for each of the above datasets by replacing
10 percent of elements in each feature, as described earlier.

@@

1. z-standardisation: $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the 
  mean and standard deviation of $\mathbf{X}$.

@@

This results in nine types of power transformation. Transformation parameters
were optimised for each feature. Subsequently the sum of residual errors 
and sum of residual errors of the central portion ($\kappa = 0.80$) were computed
for each feature after transformation. Then, each method was ranked according 
to the sum of residual errors (data without outliers) or the sum of residual
errors of the central portion (data with outliers). The average rank of each
transformation method was computed over all 100000 features in each dataset.
Average ranks for Yeo-Johnson transformations are shown in Table
\ref{tab:comparison_methods_simulations_yeo_johnson}.
Location and shift-invariant Yeo-Johnson transformation ranked best for
all datasets without outliers. The robust variant ranked best in dirty and shifted
datasets with outliers, but not in the clean dataset.
Results for Box-Cox transformations are shown in Table \ref{tab:comparison_methods_simulations_box_cox}.

@@

\begin{table}
\begin{center}
\caption{
Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 features each. The clean dataset consists of features derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations_yeo_johnson}
\begin{tabular}{l | l r r r r r r}
